---
title: "JMF_ML_dataprep"
author: "Xiaoran Sun"
date: "3/14/2023"
output: html_document
---

This is the data preparation file for the JMF special issue paper on supervsed ML.


```{r}
library(foreign)
library(psych)
setwd("~/Desktop/paper/SunProjects/Add Health")
#reading in Wave 1 In-Home Interview Data
load("Wave1InHome.rda")
wave1<-da21600.0001
spss1<-data.frame(read.spss("21600-0001-Data.sav"))
dat1<-wave1[c("AID")]
```

## Family structure
Parent presence; biological parents
```{r}
HasMom<-rep(1, 6504)
HasMom[which(as.numeric(spss1$H1RM1)==14)]<-0
HasDad<-rep(1, 6504)
HasDad[which(as.numeric(spss1$H1RF1)==14)]<-0

dat1$ResidentMotherPresence<-HasMom
dat1$ResidentFatherPresence<-HasDad

levels(spss1$H1NM1)
BioMom<-rep(NA, 6504)
BioMom[which(as.numeric(spss1$H1NM1)==4)]<-1
BioMom[which(as.numeric(spss1$H1NM1)<3)]<-0

levels(spss1$H1NF1)
BioDad<-rep(NA, 6504)
BioDad[which(as.numeric(spss1$H1NF1)==4)]<-1
BioDad[which(as.numeric(spss1$H1NF1)<3)]<-0

TwoBioParent<-rep(0, 6504)
TwoBioParent[(which(BioMom+BioDad==2))]<-1
dat1$ResidentBiologicalParents<-TwoBioParent
```

Parent married
```{r}
levels(spss1$PA10)
table(as.numeric(spss1$PA10)) #2: married
PAmarried<-rep(NA, 6504)
PAmarried[which(as.numeric(spss1$PA10)==2)]<-1
PAmarried[which(as.numeric(spss1$PA10) %in% c(1,3,4,5))]<-0
dat1$ParentMarried<-PAmarried
```

Parent partnered
```{r}
#first, check how highly correlated the partnered status is with the marital status
table(spss1$PB2)
table(as.numeric(spss1$PB2))
PApartnered<-rep(NA,6504)
PApartnered[which(as.numeric(spss1$PB2)<=2)]<-1
PApartnered[which(as.numeric(spss1$PB2)==4)]<-0
cor.test(PApartnered, PAmarried) #.78; okay
table(PApartnered, PAmarried)
dat1$ParentPartnered<-PApartnered
```


Household size
```{r}
levels(spss1$H1HR2A)
which(as.numeric(spss1$H1HR2A)==4)
levels(spss1$H1HR2E)

HHsize<-rep(NA,6504)
HHsize[which(as.numeric(spss1$H1HR2G)!=4)]<-7 #lives with seven or more
HHsize[which(as.numeric(spss1$H1HR2F)!=4 & as.numeric(spss1$H1HR2G)==4)]<-6 #lives with exactly six
HHsize[which(as.numeric(spss1$H1HR2E)!=4 & as.numeric(spss1$H1HR2F)==4)]<-5 #lives with exactly five
HHsize[which(as.numeric(spss1$H1HR2D)!=4 & as.numeric(spss1$H1HR2E)==4)]<-4
HHsize[which(as.numeric(spss1$H1HR2C)!=4 & as.numeric(spss1$H1HR2D)==4)]<-3
HHsize[which(as.numeric(spss1$H1HR2B)!=4 & as.numeric(spss1$H1HR2C)==4)]<-2
HHsize[which(as.numeric(spss1$H1HR2A)!=4 & as.numeric(spss1$H1HR2B)==4)]<-1
HHsize[which(as.numeric(spss1$H1HR2A)==4)]<-0 #lives alone
summary(as.factor(HHsize))

dat1$HouseholdSize<-HHsize
```

Number of siblings
```{r}
levels(spss1$H1HR3A)
SibNum<-rep(0,6504)
SibA<-(as.numeric(spss1$H1HR3A)==5|as.numeric(spss1$H1HR3A)==8)
SibB<-(as.numeric(spss1$H1HR3B)==5|as.numeric(spss1$H1HR3B)==8)
SibC<-(as.numeric(spss1$H1HR3C)==5|as.numeric(spss1$H1HR3C)==8)
SibD<-(as.numeric(spss1$H1HR3D)==5|as.numeric(spss1$H1HR3D)==8)
SibE<-(as.numeric(spss1$H1HR3E)==5|as.numeric(spss1$H1HR3E)==8)
SibF<-(as.numeric(spss1$H1HR3F)==5|as.numeric(spss1$H1HR3F)==8)
SibG<-(as.numeric(spss1$H1HR3G)==5|as.numeric(spss1$H1HR3G)==8)
SibH<-(as.numeric(spss1$H1HR3H)==5|as.numeric(spss1$H1HR3H)==8)
SibI<-(as.numeric(spss1$H1HR3I)==5|as.numeric(spss1$H1HR3I)==8)
SibJ<-(as.numeric(spss1$H1HR3J)==5|as.numeric(spss1$H1HR3J)==8)
SibK<-(as.numeric(spss1$H1HR3K)==5|as.numeric(spss1$H1HR3K)==8)
SibL<-(as.numeric(spss1$H1HR3L)==5|as.numeric(spss1$H1HR3L)==8)
SibM<-(as.numeric(spss1$H1HR3M)==5|as.numeric(spss1$H1HR3M)==8)
SibN<-(as.numeric(spss1$H1HR3N)==5|as.numeric(spss1$H1HR3N)==8)
SibO<-(as.numeric(spss1$H1HR3O)==5|as.numeric(spss1$H1HR3O)==8)
SibP<-(as.numeric(spss1$H1HR3P)==5|as.numeric(spss1$H1HR3P)==8)
SibNum<-SibA+SibB+SibC+SibD+SibE+SibF+SibG+SibH+SibI+SibJ+SibK+SibL+SibM+SibN+SibO+SibP
summary(SibNum)
cor(SibNum, HHsize)
head(as.numeric(spss1$H1HR3A))
dat1$NumberSiblings<-SibNum
```

Biological Birth Order
```{r}
BirthOrder<-rep(NA, 6504)
levels(spss1$H1HR14)
BirthOrder[which(as.numeric(spss1$H1HR14)==1)]<-1 #only child
levels(spss1$H1HR15)
BirthOrder[which(as.numeric(spss1$H1HR15)==1)]<-1
BirthOrder[which(as.numeric(spss1$H1HR15)==2)]<-2
BirthOrder[which(as.numeric(spss1$H1HR15)==3)]<-3
BirthOrder[which(as.numeric(spss1$H1HR15)>=4 & as.numeric(spss1$H1HR15)<=14)]<-4 #4 and above
summary(as.factor(BirthOrder))
dat1$BirthOrder<-BirthOrder
```

parent age
```{r}
head(wave1$PA2)  #responded by parents
dat1$ParentAge<-wave1$PA2
```

## Family SES

Parent education (resident parents)
* no mom/dad code as mean 
YA codings: 8 = 8th grade or less, 10 = some high school, 12 = high school graduate, 13 = vocational/technical training after high school, 14 = some college, 16 = bachelor's degree, 17 = some graduate school or some post-baccalaureate professional education, 18 = master's degree, 19 = some graduate training beyond a master’s degree, and 20 = doctoral degree or completed post-baccalaureate professional education (e.g., law school, medical school, nurse). 
```{r}
#youth report
table(wave1$H1RM1)
Ymomeduc_raw<-as.numeric(substr(wave1$H1RM1, start = 2, stop = 3)) 
Ymomeduc1<-rep(NA, 6504)
Ymomeduc1[which(Ymomeduc_raw==10)]<-0 #never went to school
Ymomeduc1[which(Ymomeduc_raw==1)]<-8 #eight grade or less
Ymomeduc1[which(Ymomeduc_raw==2)]<-10 #>8th grade/didn't graduate high school
Ymomeduc1[which(Ymomeduc_raw==3)]<-11 #>Business/trade/voc. school instead high school 
Ymomeduc1[which(Ymomeduc_raw==4|Ymomeduc_raw==5)]<-12 #High school graduate; GED 
Ymomeduc1[which(Ymomeduc_raw==6)]<-13 #Business/trade/voc. school after high school
Ymomeduc1[which(Ymomeduc_raw==7)]<-14 #College/didn't graduate
Ymomeduc1[which(Ymomeduc_raw==8)]<-16 #Graduated from college/university
Ymomeduc1[which(Ymomeduc_raw==9)]<-17 #Prof training beyond 4-year college/university
table(Ymomeduc1)
Ymomeduc1[which(Ymomeduc_raw==11|Ymomeduc_raw==12)]<-NA
mean(Ymomeduc1, na.rm = TRUE)
length(spss1[as.numeric(spss1$H1RM1)==14,]$AID) #legitimate skip
Ymomeduc1[which(as.numeric(spss1$H1RM1)==14)]<- 13.22558 #no resident mom
summary(as.factor(Ymomeduc1))

table(wave1$H1RF1)
Ydadeduc_raw<-as.numeric(substr(wave1$H1RF1, start = 2, stop = 3)) 
Ydadeduc1<-rep(NA, 6504)
Ydadeduc1[which(Ydadeduc_raw==10)]<-0 #never went to school
Ydadeduc1[which(Ydadeduc_raw==1)]<-8 #eight grade or less
Ydadeduc1[which(Ydadeduc_raw==2)]<-10 #>8th grade/didn't graduate high school
Ydadeduc1[which(Ydadeduc_raw==3)]<-11 #>Business/trade/voc. school instead high school 
Ydadeduc1[which(Ydadeduc_raw==4|Ydadeduc_raw==5)]<-12 #High school graduate; GED 
Ydadeduc1[which(Ydadeduc_raw==6)]<-13 #Business/trade/voc. school after high school
Ydadeduc1[which(Ydadeduc_raw==7)]<-14 #College/didn't graduate
Ydadeduc1[which(Ydadeduc_raw==8)]<-16 #Graduated from college/university
Ydadeduc1[which(Ydadeduc_raw==9)]<-17 #Prof training beyond 4-year college/university
table(Ydadeduc1)
Ydadeduc1[which(Ydadeduc_raw==11|Ydadeduc_raw==12)]<-NA
mean(Ydadeduc1, na.rm = TRUE)
length(spss1[as.numeric(spss1$H1RF1)==14,]$AID) #legitimate skip
Ydadeduc1[which(as.numeric(spss1$H1RF1)==14)]<- 13.36908 #no resident dad
summary(as.factor(Ydadeduc1))

dat1$MotherEducation<-Ymomeduc1
dat1$FatherEducation<-Ydadeduc1
```

Parent occupation
* recode into five ordered categories (Feliciano & Lanuza, 2017):
  professionals, office/sales workers, blue-collar workers, crafts/military/farm/other, and unemployed (5 to 1)
* Note: No mom/dad code as mean
```{r}
#youth report
levels(wave1$H1RM4)
Ymomjob<-as.numeric(substr(wave1$H1RM4, start = 2, stop = 3))
Ymomjob_re<-rep(NA, 6504)
Ymomjob_re[which(Ymomjob<=4)]<-5  
Ymomjob_re[which(Ymomjob==5|Ymomjob==6)]<-4
Ymomjob_re[which(Ymomjob==7|Ymomjob==9|Ymomjob==10|Ymomjob==11|Ymomjob==12)]<-3
Ymomjob_re[which(Ymomjob==8|Ymomjob==13|Ymomjob==14|Ymomjob==15)]<-2
Ymomjob_re[which(Ymomjob==16)]<-1
mean(Ymomjob_re, na.rm = TRUE)
Ymomjob_re[which(as.numeric(spss1$H1RM4)==18)]<-3.350598
summary(as.factor(Ymomjob_re))

Ydadjob<-as.numeric(substr(wave1$H1RF4, start = 2, stop = 3))
Ydadjob_re<-rep(NA, 6504)
Ydadjob_re[which(Ydadjob<=4)]<-5  
Ydadjob_re[which(Ydadjob==5|Ydadjob==6)]<-4
Ydadjob_re[which(Ydadjob==7|Ydadjob==9|Ydadjob==10|Ydadjob==11|Ydadjob==12)]<-3
Ydadjob_re[which(Ydadjob==8|Ydadjob==13|Ydadjob==14|Ydadjob==15)]<-2
Ydadjob_re[which(Ydadjob==16)]<-1
mean(Ydadjob_re, na.rm = TRUE)
Ydadjob_re[which(as.numeric(spss1$H1RF4)==18)]<-3.347201
summary(as.factor(Ydadjob_re))

dat1$MotherOccupationPrestige<-Ymomjob_re
dat1$FatherOccupationPrestige<-Ydadjob_re
```

Parent working for pay
0= no or None for occupation; 1 = yes; legitimate skip = mean
```{r}
table(wave1$H1RM5)
Ymompay<-as.numeric(substr(wave1$H1RM5, start = 2, stop = 2))
table(Ymompay)
Ymompay[which(Ymomjob==16)]<-0
table(Ymompay)
mean(Ymompay, na.rm = TRUE)
Ymompay[which(as.numeric(spss1$H1RM4)==18)]<-0.7734247
table(Ymompay)

table(wave1$H1RF5)
Ydadpay<-as.numeric(substr(wave1$H1RF5, start = 2, stop = 2))
table(Ydadpay)
Ydadpay[which(Ydadjob==16)]<-0
table(Ydadpay)
mean(Ydadpay, na.rm = TRUE)
Ydadpay[which(as.numeric(spss1$H1RF4)==18)]<-0.929593
table(Ydadpay)

dat1$MotherWorkForPay<-Ymompay
dat1$FatherWorkForPay<-Ydadpay

```

family income (in thousand, log)
```{r}
dat1$FamilyIncome<-log(wave1$PA55+1)
exp(3.17)-1
```

parent receiving public assistance
```{r}
PAassistance<-as.numeric(substr(wave1$PA21, start = 2, stop = 2))
dat1$PublicAssistance<-PAassistance
```

Family receive welfare
```{r}
welfare1<-as.numeric(substr(wave1$PA57B, start = 2, stop = 2))
welfare2<-as.numeric(substr(wave1$PA57C, start = 2, stop = 2))
welfare3<-as.numeric(substr(wave1$PA57D, start = 2, stop = 2))
welfare<-rowSums(cbind(welfare1, welfare2, welfare3), na.rm = TRUE)
welfare[which(is.na(welfare1)&is.na(welfare2)&is.na(welfare3))]<-NA
welfare[which(welfare>=1)]<-1
dat1$Welfare<-welfare
```

Parent economic hardship;
```{r}
levels(spss1$PA56)
PAecohard<-1-as.numeric(substr(wave1$PA56, start = 2, stop = 2))
dat1$EconomicHardship<-PAecohard
```

## family relationships and parenting
family social support (Ryabov, 2016)
```{r}
levels(wave1$H1PR3)
levels(wave1$H1PR5)
levels(wave1$H1PR7)
levels(wave1$H1PR8)
famsup1<-as.numeric(substr(wave1$H1PR3, start = 2, stop = 2))
famsup1[which(famsup1==6)]<-NA #"does not apply"
famsup2<-as.numeric(substr(wave1$H1PR5, start = 2, stop = 2))
famsup2[which(famsup2==6)]<-NA #"does not apply"
famsup3<-as.numeric(substr(wave1$H1PR7, start = 2, stop = 2))
famsup3[which(famsup3==6)]<-NA #"does not apply"
famsup4<-as.numeric(substr(wave1$H1PR8, start = 2, stop = 2))
famsup4[which(famsup4==6)]<-NA #"does not apply"

famsup<-rowMeans(cbind(famsup1, famsup2, famsup3, famsup4), na.rm = TRUE)
dat1$FamilySocialSupport<-famsup
```

Shared dinner with parents
*no parent = mean
```{r}
levels(spss1$H1WP8)
dinner<-as.numeric(substr(wave1$H1WP8, start = 2, stop = 2)) 
mean(dinner, na.rm=TRUE)
dinner[which(as.numeric(spss1$H1WP8)==10)]<-4.608819
dat1$SharedDinner<-dinner
```

intergenerational closure
```{r}
levels(wave1$PC17)
PAclosure<-as.numeric(substr(wave1$PC17, start = 2, stop = 2))
dat1$IntergenerationalClosure<-PAclosure
```

Parent connection to child's friends
```{r}
levels(wave1$PC10)
PAconnection1<-as.numeric(substr(wave1$PC10, start = 2, stop = 2))
PAconnection1[which(PAconnection1==2)]<-1 #doesn't go to school-- also know
table(PAconnection1)

table(wave1$PC11)
PAconnection2<-as.numeric(substr(wave1$PC11, start = 2, stop = 2))
table(PAconnection2)

table(wave1$PC12)
PAconnection3<-as.numeric(substr(wave1$PC12, start = 2, stop = 2))
table(PAconnection3)

table(spss1$PC13) #recode: bad & neither good/bad & don't know ->0; good influence-> 1
PAconnection4<-as.numeric(spss1$PC13)
table(PAconnection4)
PAconnection4[which(PAconnection4==4)]<-NA #refused to answer
table(PAconnection4)
PAconnection4[which(PAconnection4!=3)]<-0
table(PAconnection4)

PAconnection<-PAconnection1+PAconnection2+PAconnection3+PAconnection4

dat1$ParentConnectionFriends<-PAconnection
```

Mother/father-adolescent relationship quality
```{r}
mrel1<-as.numeric(substr(wave1$H1WP9, start = 2, stop = 2))
mrel2<-as.numeric(substr(wave1$H1WP10, start = 2, stop = 2))
mrel3<-6-as.numeric(substr(wave1$H1PF1, start = 2, stop = 2))
mrel4<-6-as.numeric(substr(wave1$H1PF4, start = 2, stop = 2))
mrel5<-6-as.numeric(substr(wave1$H1PF5, start = 2, stop = 2))
mrel<-rowMeans(cbind(mrel1, mrel2, mrel3, mrel4, mrel5), na.rm = TRUE)
mean(mrel, na.rm=TRUE)
mrel[which(as.numeric(spss1$H1WP9)==7)]<-4.430109 #no mom

drel1<-as.numeric(substr(wave1$H1WP13, start = 2, stop = 2))
drel2<-as.numeric(substr(wave1$H1WP14, start = 2, stop = 2))
drel3<-6-as.numeric(substr(wave1$H1PF23, start = 2, stop = 2))
drel4<-6-as.numeric(substr(wave1$H1PF24, start = 2, stop = 2))
drel5<-6-as.numeric(substr(wave1$H1PF25, start = 2, stop = 2))
drel<-rowMeans(cbind(drel1, drel2, drel3, drel4, drel5), na.rm = TRUE)
mean(drel, na.rm=TRUE)
drel[which(as.numeric(spss1$H1WP13)==7)]<-4.243704

dat1$MotherRelationshipQuality<-mrel
dat1$FatherRelationshipQuality<-drel
```

shared activities (in general)
* no mom/dad = mean
```{r}
mtime1<-as.numeric(substr(wave1$H1WP17A, start = 2, stop = 2))
mtime2<-as.numeric(substr(wave1$H1WP17B, start = 2, stop = 2))
mtime3<-as.numeric(substr(wave1$H1WP17C, start = 2, stop = 2))
mtime4<-as.numeric(substr(wave1$H1WP17D, start = 2, stop = 2))
mtime5<-as.numeric(substr(wave1$H1WP17E, start = 2, stop = 2))
mtime6<-as.numeric(substr(wave1$H1WP17F, start = 2, stop = 2))
mtime7<-as.numeric(substr(wave1$H1WP17G, start = 2, stop = 2))
mtime8<-as.numeric(substr(wave1$H1WP17H, start = 2, stop = 2))
mtime9<-as.numeric(substr(wave1$H1WP17I, start = 2, stop = 2))
mtime10<-as.numeric(substr(wave1$H1WP17J, start = 2, stop = 2))
mtime<-mtime1+mtime2+mtime3+mtime4+mtime5+mtime6+mtime7+mtime8+mtime9+mtime10
mean(mtime, na.rm = TRUE)
mtime[which(as.numeric(spss1$H1WP17A)==4)]<- 3.951984 #no mom
summary(as.factor(mtime))

dtime1<-as.numeric(substr(wave1$H1WP18A, start = 2, stop = 2))
dtime2<-as.numeric(substr(wave1$H1WP18B, start = 2, stop = 2))
dtime3<-as.numeric(substr(wave1$H1WP18C, start = 2, stop = 2))
dtime4<-as.numeric(substr(wave1$H1WP18D, start = 2, stop = 2))
dtime5<-as.numeric(substr(wave1$H1WP18E, start = 2, stop = 2))
dtime6<-as.numeric(substr(wave1$H1WP18F, start = 2, stop = 2))
dtime7<-as.numeric(substr(wave1$H1WP18G, start = 2, stop = 2))
dtime8<-as.numeric(substr(wave1$H1WP18H, start = 2, stop = 2))
dtime9<-as.numeric(substr(wave1$H1WP18I, start = 2, stop = 2))
dtime10<-as.numeric(substr(wave1$H1WP18J, start = 2, stop = 2))
dtime<-dtime1+dtime2+dtime3+dtime4+dtime5+dtime6+dtime7+dtime8+dtime9+dtime10
mean(dtime, na.rm = TRUE)
dtime[which(as.numeric(spss1$H1WP18A)==4)]<- 2.910392 #no dad
summary(as.factor(dtime))

dat1$MotherSharedActivity<-mtime
dat1$FatherSharedActivity<-dtime
```

Parent relationship quality with spouse/partner
* Skip pattern: PApartnered = 0: mean imputation
* three items not at the same scale / order; need to reverse code & standardize; the reference (King et al., 2020) isn't clear about how they computed the composite score
```{r}
table(wave1$PB18)
PArelation1<-scale(wave1$PB18)
table(wave1$PB19)
PArelation2<-scale(1-as.numeric(substr(wave1$PB19, start = 2, stop = 2)))
table(PArelation2)
table(wave1$PB20)
PArelation3<-scale(as.numeric(substr(wave1$PB20, start = 2, stop = 2)))
table(PArelation3)

PArelation<-rowMeans(cbind(PArelation1, PArelation2, PArelation3), na.rm = T)
mean(PArelation, na.rm = T)

PArelation[which(PApartnered==0)]<--0.0005413742
dat1$ParentPartnerRelationship<-PArelation
```

Parental control
* reverse-coded
*no parent = mean
```{r}
control1<-1-as.numeric(substr(wave1$H1WP1, start = 2, stop = 2))
control2<-1-as.numeric(substr(wave1$H1WP2, start = 2, stop = 2))
control3<-1-as.numeric(substr(wave1$H1WP3, start = 2, stop = 2))
control4<-1-as.numeric(substr(wave1$H1WP4, start = 2, stop = 2))
control5<-1-as.numeric(substr(wave1$H1WP5, start = 2, stop = 2))
control6<-1-as.numeric(substr(wave1$H1WP6, start = 2, stop = 2))
control7<-1-as.numeric(substr(wave1$H1WP7, start = 2, stop = 2))
control<-rowSums(cbind(control1, control2, control3, control4, control5, control6, control7), na.rm = TRUE) 
mean(control, na.rm=TRUE)
control[which(as.numeric(spss1$H1WP1)==4)]<-1.807349
summary(as.factor(control))
dat1$ParentControl<-control
```

mother/father supervision
* no mom/dad = mean
```{r}
#most/all of the time/takes = 1; other = 0
levels(wave1$H1RM11)
mspv1<-as.numeric(substr(wave1$H1RM11, start = 2, stop = 2))
mspv1[which(mspv1==1|mspv1==2|mspv1==6)]<-1
mspv1[which(mspv1==3|mspv1==4|mspv1==5)]<-0
mspv2<-as.numeric(substr(wave1$H1RM12, start = 2, stop = 2))
mspv2[which(mspv2==1|mspv2==2|mspv2==6)]<-1
mspv2[which(mspv2==3|mspv2==4|mspv2==5)]<-0
mspv3<-as.numeric(substr(wave1$H1RM13, start = 2, stop = 2))
mspv3[which(mspv3==1|mspv3==2)]<-1
mspv3[which(mspv3==3|mspv3==4|mspv3==5)]<-0
mspv<-rowSums(cbind(mspv1, mspv2, mspv3), na.rm = TRUE)
mean(mspv)
mspv[which(as.numeric(spss1$H1RM11)==8)]<-2.090406

dspv1<-as.numeric(substr(wave1$H1RF11, start = 2, stop = 2))
dspv1[which(dspv1==1|dspv1==2|dspv1==6)]<-1
dspv1[which(dspv1==3|dspv1==4|dspv1==5)]<-0
dspv2<-as.numeric(substr(wave1$H1RF12, start = 2, stop = 2))
dspv2[which(dspv2==1|dspv2==2|dspv2==6)]<-1
dspv2[which(dspv2==3|dspv2==4|dspv2==5)]<-0
dspv3<-as.numeric(substr(wave1$H1RF13, start = 2, stop = 2))
dspv3[which(dspv3==1|dspv3==2)]<-1
dspv3[which(dspv3==3|dspv3==4|dspv3==5)]<-0
dspv<-rowSums(cbind(dspv1, dspv2, dspv3), na.rm = TRUE)
mean(dspv)
dspv[which(as.numeric(spss1$H1RF11)==8)]<-1.100092

dat1$MotherSupervision<-mspv
dat1$FatherSupervision<-dspv
```
## parent involvement with education
parental involvement with schoolwork
* no mom/dad = mean
```{r}
minvolve1<-as.numeric(substr(wave1$H1WP17H, start = 2, stop = 2))
minvolve2<-as.numeric(substr(wave1$H1WP17I, start = 2, stop = 2))
minvolve3<-as.numeric(substr(wave1$H1WP17J, start = 2, stop = 2))
minvolve<-minvolve1+minvolve2+minvolve3
mean(minvolve, na.rm = TRUE)
minvolve[which(as.numeric(spss1$H1WP17H)==4)]<-1.280745 #no mom
summary(as.factor(minvolve))

dinvolve1<-as.numeric(substr(wave1$H1WP18H, start = 2, stop = 2))
dinvolve2<-as.numeric(substr(wave1$H1WP18I, start = 2, stop = 2))
dinvolve3<-as.numeric(substr(wave1$H1WP18J, start = 2, stop = 2))
dinvolve<-dinvolve1+dinvolve2+dinvolve3
mean(dinvolve, na.rm = TRUE)
dinvolve[which(as.numeric(spss1$H1WP18H)==4)]<- 1.068472 #no dad
summary(as.factor(dinvolve))

dat1$MotherInvolveSchool<-minvolve
dat1$FatherInvolveSchool<-dinvolve
```

parent involvement in PTA
```{r}
levels(wave1$PA27A)
PAPTA<-as.numeric(substr(wave1$PA27A, start = 2, stop = 2))
dat1$ParentPTA<-PAPTA
```

Parent in school fund-raising
```{r}
levels(spss1$PC28)
PAfund<-as.numeric(substr(wave1$PC28, start = 2, stop = 2)) #many missing due to legitimate skip -- adolescent not in school now; for this project because the focus is on family variables (not school variables), this just is treated as missing
dat1$ParentFundRaise<-PAfund
```

Parent met teachers
```{r}
levels(wave1$PC27)
PAteacher<-as.numeric(substr(wave1$PC27, start = 2, stop = 2))#many missing due to legitimate skip -- adolescent not in school now; for this project because the focus is on family variables (not school variables), this just is treated as missing
dat1$ParentMetTeacher<-PAteacher
```

Adolescents' perceptions of parents' educational expectations
* no mom/dad = mean
```{r}
mexp<-rowMeans(cbind(wave1$H1WP11, wave1$H1WP12), na.rm = TRUE)
mean(mexp, na.rm = TRUE)
mexp[which(as.numeric(spss1$H1WP11)==7)]<-4.341575 #no mom->0
dexp<-rowMeans(cbind(wave1$H1WP15, wave1$H1WP16), na.rm = TRUE)
mean(dexp, na.rm = TRUE)
dexp[which(as.numeric(spss1$H1WP15)==7)]<-4.363035 #no dad->0

summary(dexp)

dat1$MotherEducationExpectation<-mexp
dat1$FatherEducationExpectation<-dexp
```
## family sociocultural characteristics
Parent nativity
*No mom/dad code as mean
```{r}
#youth report
YmomUSborn<-as.numeric(substr(wave1$H1RM2, start = 2, stop = 2))
mean(YmomUSborn, na.rm = TRUE)
YmomUSborn[which(as.numeric(spss1$H1RM2)==4)]<- 0.8798627
summary(as.factor(YmomUSborn))

YdadUSborn<-as.numeric(substr(wave1$H1RF2, start = 2, stop = 2))
mean(YdadUSborn, na.rm = TRUE)
YdadUSborn[which(as.numeric(spss1$H1RF2)==4)]<- 0.8700154
summary(as.factor(YdadUSborn))

dat1$MotherNativity<-YmomUSborn
dat1$FatherNativity<-YdadUSborn
```


Parent years in the U.S.
```{r}
summary(spss1$PC65_Y)
table(as.numeric(spss1$PC65_Y))
PAyearsUS<-(as.numeric(wave1$IYEAR)+1993)- (as.numeric(substr(wave1$PC65_Y, start = 2, stop = 3))+1976) +((as.numeric(substr(wave1$IMONTH, start=2, stop=3)))-(as.numeric(substr(wave1$PC65_M, start=2, stop=3))))/12
PAyearsUS[which(as.numeric(spss1$PC65_Y)==20)] <- dat1[which(as.numeric(spss1$PC65_Y)==20),]$ParentAge #legitimate skip; but don't need to control bc of filling in age
summary(PAyearsUS)
dat1$ParentYearsInUS<-PAyearsUS
#forgot to add in the first round of data curation; add this column back to where it belongs
#library(tibble)
#dat1_c<-add_column(dat1, ParentYearsInUS = PAyearsUS, .after = 'FatherNativity')
#dat1<-dat1_c
```

Parent religiosity
* no religion--coded as lowest (Stokes, 2008)
* reverse-coded
```{r}
levels(wave1$PA23)
levels(spss1$PA23)
PArelig1<-5-as.numeric(substr(wave1$PA23, start = 2, stop = 2))
PArelig1[which(as.numeric(spss1$PA23)==6)]<-1
PArelig2<-5-as.numeric(substr(wave1$PA24, start = 2, stop = 2))
PArelig2[which(as.numeric(spss1$PA24)==6)]<-1
PArelig<-rowMeans(cbind(PArelig1, PArelig2), na.rm = TRUE)
dat1$ParentReligiosity<-PArelig
hist(PArelig)
```



English as home language
```{r}
EnglishHome<-as.numeric(substr(wave1$H1GI10, start = 2, stop = 2))
EnglishHome[which(EnglishHome==2|EnglishHome==3)]<-0
dat1$EnglishHome<-EnglishHome
```

## family health resources & behaviors
parent self-rated health
* reverse-coded
```{r}
levels(wave1$PA58)
PAhealth<-6-as.numeric(substr(wave1$PA58, start = 2, stop = 2))
dat1$ParentHealth<-PAhealth
```

parent smoke
```{r}
levels(wave1$PA64)
PAsmoke<-as.numeric(substr(wave1$PA64, start = 2, stop = 2))
PAsmoke[which(as.numeric(spss1$PA64)==4)]<-0 #legitimate skip-> no smokers in family
summary(as.factor(PAsmoke))
dat1$ParentSmoke<-PAsmoke
```

Parent alcoholic (household)
```{r}
BioMomAlc<-as.numeric(substr(wave1$PC49E_2, start = 2, stop = 2))
BioDadAlc<-as.numeric(substr(wave1$PC49E_3, start = 2, stop = 2))
malcohol<-rep(NA, 6504)
malcohol[(which(BioMom==1 & BioMomAlc==1))]<-1
malcohol[(which(BioMom==1 & BioMomAlc==0))]<-0
mean(malcohol, na.rm=TRUE)
malcohol[which(HasMom==0)]<-0.0141125
summary(as.factor(malcohol))

dalcohol<-rep(NA, 6504)
dalcohol[(which(BioDad==1 & BioDadAlc==1))]<-1
dalcohol[(which(BioDad==1 & BioDadAlc==0))]<-0
mean(dalcohol, na.rm=TRUE)
dalcohol[which(HasDad==0)]<-0.06048632
summary(as.factor(dalcohol))

dat1$MotherAlcoholic<-malcohol
dat1$FatherAlcoholic<-dalcohol
```

Parent obese (household)
```{r}
BioMomObese<-as.numeric(substr(wave1$PC49A_2, start = 2, stop = 2))
BioDadObese<-as.numeric(substr(wave1$PC49A_3, start = 2, stop = 2))
MomObese<-rep(NA, 6504)
MomObese[(which(BioMom==1 & BioMomObese==1))]<-1
MomObese[(which(BioMom==1 & BioMomObese==0))]<-0
mean(MomObese, na.rm=TRUE)
MomObese[which(HasMom==0)]<-0.1901437
summary(as.factor(MomObese))

DadObese<-rep(NA, 6504)
DadObese[(which(BioDad==1 & BioDadObese==1))]<-1
DadObese[(which(BioDad==1 & BioDadObese==0))]<-0
mean(DadObese, na.rm=TRUE)
DadObese[which(HasDad==0)]<-0.1127865
summary(as.factor(DadObese))

dat1$MotherObese<-MomObese
dat1$FatherObese<-DadObese
```

Mother/father disabled
*No mom/dad code as mean
```{r}
mdisable<-as.numeric(substr(wave1$H1RM10, start = 2, stop = 2))
mean(mdisable, na.rm = TRUE)
mdisable[which(as.numeric(spss1$H1RM10)==4)]<- 0.04313021
summary(as.factor(mdisable))

ddisable<-as.numeric(substr(wave1$H1RF10, start = 2, stop = 2))
mean(ddisable, na.rm = TRUE)
ddisable[which(as.numeric(spss1$H1RF10)==4)]<- 0.05724351
summary(as.factor(ddisable))

dat1$MotherDisable<-mdisable
dat1$FatherDisable<-ddisable
```

Smoker(s) in household
```{r}
HHsmoke<-as.numeric(substr(wave1$PA63, start = 2, stop = 2))
dat1$SmokerHousehold<-HHsmoke
```

Illegal drugs in household
```{r}
HHdrug<-as.numeric(substr(wave1$H1TO52, start = 2, stop = 2))
dat1$DrugHousehold<-HHdrug
```

Family access to medical care (reverse-coded)
```{r}
fammed<-5-as.numeric(substr(wave1$PA59, start = 2, stop = 2))
summary(as.factor(fammed))
dat1$FamilyAccessMedicalCare<-fammed
```

## parent and family relationship history
Parents' number of previous marriage or marriage-like relationships
```{r}
summary(as.factor(wave1$PA40))
PArelationnum<-as.numeric(substr(wave1$PA40, start = 2, stop = 2))
summary(as.factor(PArelationnum))
dat1$ParentPreviousRelationship<-PArelationnum
```

Nonresident biological father unknown or deceased
* legitimate skip: dad is biological father
```{r}
table(spss1$H1NF1)
table(as.numeric(spss1$H1NF1))
table(spss1$H1NF2)
table(as.numeric(spss1$H1NF2))
nrbiodad<-rep(NA, 6504)
nrbiodad[which(as.numeric(spss1$H1NF1)==1|as.numeric(spss1$H1NF1)==5)]<-0
nrbiodad[which(as.numeric(spss1$H1NF2)==1)]<-0
nrbiodad[which(as.numeric(spss1$H1NF1)==2 & as.numeric(spss1$H1NF2)==2)]<-1
mean(nrbiodad, na.rm=T)

nrbiodad[which(as.numeric(spss1$H1NF1)==4)]<-0.7259314
summary(as.factor(nrbiodad))

dat1$NonresidentBioFatherDeceased<-nrbiodad
```

Number of years child had lived with stepfather
* legitimate skip: no stepfather in the house roster
```{r}
table(spss1$H1HR6A)
stepdadA<-as.numeric(as.numeric(spss1$H1HR6A)==2)
stepdadB<-as.numeric(as.numeric(spss1$H1HR6B)==2)
stepdadC<-as.numeric(as.numeric(spss1$H1HR6C)==2)
stepdadD<-as.numeric(as.numeric(spss1$H1HR6D)==2)
stepdadE<-as.numeric(as.numeric(spss1$H1HR6E)==2)
stepdadF<-as.numeric(as.numeric(spss1$H1HR6F)==2)
stepdadG<-as.numeric(as.numeric(spss1$H1HR6G)==2)
stepdadH<-as.numeric(as.numeric(spss1$H1HR6H)==2)
stepdadI<-as.numeric(as.numeric(spss1$H1HR6I)==2)
stepdadJ<-as.numeric(as.numeric(spss1$H1HR6J)==2)

stepdadtot<-stepdadA+stepdadB+stepdadC+stepdadD+stepdadE+stepdadF+stepdadG+stepdadH+stepdadI+stepdadJ
table(stepdadtot) #510 stepdad; at most 1 stepdad a family

stepdaddat<-data.frame(cbind(stepdadA, stepdadB, stepdadC, stepdadD, stepdadE, stepdadF, stepdadG, stepdadH, stepdadI, stepdadJ))

summary(wave1$H1HR9A)
summary(wave1$H1HR10A)

alwayssameHHA<-as.numeric(substr(wave1$H1HR9A, start = 2, stop = 2))
alwayssameHHB<-as.numeric(substr(wave1$H1HR9B, start = 2, stop = 2))
alwayssameHHC<-as.numeric(substr(wave1$H1HR9C, start = 2, stop = 2))
alwayssameHHD<-as.numeric(substr(wave1$H1HR9D, start = 2, stop = 2))
alwayssameHHE<-as.numeric(substr(wave1$H1HR9E, start = 2, stop = 2))
alwayssameHHF<-as.numeric(substr(wave1$H1HR9F, start = 2, stop = 2))
alwayssameHHG<-as.numeric(substr(wave1$H1HR9G, start = 2, stop = 2))
alwayssameHHH<-as.numeric(substr(wave1$H1HR9H, start = 2, stop = 2))
alwayssameHHI<-as.numeric(substr(wave1$H1HR9I, start = 2, stop = 2))
alwayssameHHJ<-as.numeric(substr(wave1$H1HR9J, start = 2, stop = 2))

alwayssamedat<-data.frame(cbind(alwayssameHHA, alwayssameHHB, alwayssameHHC, alwayssameHHD, alwayssameHHE, alwayssameHHF, alwayssameHHG, alwayssameHHH, alwayssameHHI, alwayssameHHJ))

sameHHyears<-data.frame(cbind(wave1$H1HR10A, wave1$H1HR10B, wave1$H1HR10C, wave1$H1HR10D, wave1$H1HR10E, wave1$H1HR10F, wave1$H1HR10G, wave1$H1HR10H, wave1$H1HR10I, wave1$H1HR10J))
Yage<-(as.numeric(wave1$IYEAR)+1993)- (as.numeric(wave1$H1GI1Y)+1973)+((as.numeric(substr(wave1$IMONTH, start=2, stop=3)))-(as.numeric(substr(wave1$H1GI1M, start=2, stop=3))))/12
stepdadyears<-rep(NA, 6504)
for(i in 1:6504){
  for (j in 1:10){
    if(stepdaddat[i,j]==1){
      if(!is.na(alwayssamedat[i,j]) & alwayssamedat[i,j]==1){
        stepdadyears[i]<- Yage[i]
      }
      if(!is.na(alwayssamedat[i,j] ) & alwayssamedat[i,j]==0){
        stepdadyears[i]<-sameHHyears[i,j]
      }
    }
  }
}

describe(stepdadyears)
mean(stepdadyears, na.rm = T)
summary(stepdadyears[stepdadtot==0])

stepdadyears[stepdadtot==0]<-7.569198

dat1$YearsLiveStepfather<-stepdadyears
```

Child born to adolescent parent
* should only count the biological parent case
```{r}
Yage<-(as.numeric(wave1$IYEAR)+1993)- (as.numeric(wave1$H1GI1Y)+1973)+((as.numeric(substr(wave1$IMONTH, start=2, stop=3)))-(as.numeric(substr(wave1$H1GI1M, start=2, stop=3))))/12
summary(wave1$PA2)
PAbornage<-wave1$PA2-Yage
PCrelat<-as.numeric(substr(wave1$PC1, start = 2, stop = 3))
table(PCrelat)
PAbornage[which(PCrelat!=1 & PCrelat!=9)]<-NA
summary(PAbornage)
PAbornage[which(PAbornage<12)]<-NA
PAteen<-as.numeric(PAbornage<18)
summary(as.factor(PAteen))
dat1$BornToAdolescentParent<-PAteen
```

## Adverse family experiences

Load wave 4 data
```{r}
load("Wave4InHome.rda")
wave4<-da21600.0022
dat4<-wave4[c("AID")]
```

Emotional abuse
```{r}
table(wave4$H4MA1)
emoabuse<-as.numeric(substr(wave4$H4MA1, start = 2, stop = 2))
table(emoabuse)
emoabuse[which(emoabuse==6)]<-0
dat4$EmotionalAbuse<-emoabuse
```

Physical abuse
```{r}
table(wave4$H4MA3)
phyabuse<-as.numeric(substr(wave4$H4MA3, start = 2, stop = 2))
table(phyabuse)
phyabuse[which(phyabuse==6)]<-0
dat4$PhysicalAbuse<-phyabuse
```

Biological mother incarceration times
```{r}
summary(wave4$H4WP3)
summary(wave4$H4WP4)
momincar<-as.numeric(substr(wave4$H4WP3, start = 2, stop = 2))
momincart<-as.numeric(substr(wave4$H4WP4, start = 2, stop = 2))
momincart[which(momincar==0)]<-0
summary(as.factor(momincart))
dat4$BioMotherIncarceration<-momincart
```

Biological father incarceration times
```{r}
summary(wave4$H4WP9)
summary(wave4$H4WP10)
dadincar<-as.numeric(substr(wave4$H4WP9, start = 2, stop = 2))
table(dadincar)
dadincart<-wave4$H4WP10
table(dadincart)
dadincart[which(dadincar==0)]<-0
summary(as.factor(dadincart))
dat4$BioFatherIncarceration<-dadincart
```

## Adolescent demographic characteristics
Biological Sex, Age
```{r}
dat1$Sex<-as.numeric(substr(wave1$BIO_SEX, start=2, stop = 2))-1
wave1$Age<-(as.numeric(wave1$IYEAR)+1993)- (as.numeric(wave1$H1GI1Y)+1973)+((as.numeric(substr(wave1$IMONTH, start=2, stop=3)))-(as.numeric(substr(wave1$H1GI1M, start=2, stop=3))))/12
dat1$Age<-wave1$Age
describe(dat1$Age)
```

Race/Ethnicity
```{r}
#Latino
latino<-as.numeric(substr(wave1$H1GI4, start = 2, stop = 2))
#White
white<-as.numeric(substr(wave1$H1GI6A, start = 2, stop = 2))
#African American
AA<-as.numeric(substr(wave1$H1GI6B, start = 2, stop = 2))
#Native American
Native<-as.numeric(substr(wave1$H1GI6C, start = 2, stop = 2))
#Asian
Asian<-as.numeric(substr(wave1$H1GI6D, start = 2, stop = 2))
#Other
Other<-as.numeric(substr(wave1$H1GI6E, start = 2, stop = 2))

#Using white as a reference
dat1$Latinx<-latino
dat1$Black<-AA
dat1$NativeAmerican<-Native
dat1$AAPI<-Asian
dat1$OtherRace<-Other
```

## Educational attainment at Wave 4
Recode: 8 = 8th grade or less, 10 = some high school, 12 = high school graduate, 13 = vocational/technical training after high school, 14 = some college, 16 = bachelor's degree, 17 = some graduate school or some post-baccalaureate professional education, 18 = master's degree, 19 = some graduate training beyond a master’s degree, and 20 = doctoral degree or completed post-baccalaureate professional education (e.g., law school, medical school, nurse).
```{r}
education<-as.numeric(substr(wave4$H4ED2, start = 2, stop = 3)) #highest level of education
edu_re<-rep(NA, 5114)
edu_re[which(education==1)]<-8
edu_re[which(education==2)]<-10
edu_re[which(education==3)]<-12
edu_re[which(education==4|education==5)]<-13
edu_re[which(education==6)]<-14
edu_re[which(education==7)]<-16
edu_re[which(education==8|education==12)]<-17
edu_re[which(education==9)]<-18
edu_re[which(education==10)]<-19
edu_re[which(education==11|education==13)]<-20
summary(as.factor(edu_re))

dat4$EducationAttainment<-edu_re
```

Combine dat1 & dat4
```{r}
dat14<-merge(dat1, dat4, by = "AID")
write.csv(dat14, "dat14_JMF.csv", row.names = F)
```

## data preparation
Check missingness
```{r}
colmiss<-sapply(dat14, function(x) sum(is.na (x)))/51.14 #variable-wise
summary(colmiss) 
which(colmiss>=25)
dat14_re<-dat14[,c(1:35, 38:71)] #delete these two variables
colmiss_re<-sapply(dat14_re, function(x) sum(is.na (x)))/51.14 #variable-wise
summary(colmiss_re) #range & descriptives of missingness
```

corelation checks
```{r}

for(i in 2:67){
  for (j in (i+1):68){
    r = cor(dat14_re[,i], dat14_re[,j], use="pairwise.complete.obs")
    if (r > .50){
      print(paste(r, names(dat14_re)[i], names(dat14_re)[j]))
    }
  }
}

#none were correlated more than .80

```

MissForest imputation
```{r}
library(missForest)
doParallel::registerDoParallel(cores = 4) 
doRNG::registerDoRNG(seed = 123)
#need to specify the categorical variables
names(dat14_re)
factorcols<-c("ResidentMotherPresence", "ResidentFatherPresence", "ResidentBiologicalParents", "ParentMarried", "BirthOrder", "MotherWorkForPay", "FatherWorkForPay", "PublicAssistance", "Welfare", "EconomicHardship", "ParentPartnered","ParentPTA", "MotherNativity", "FatherNativity", "EnglishHome", "ParentSmoke", "MotherAlcoholic", "FatherAlcoholic", "MotherObese", "FatherObese", "MotherDisable", "FatherDisable", "SmokerHousehold", "DrugHousehold", "NonresidentBioFatherDeceased", "BornToAdolescentParent", "Sex", "Latinx", "Black" ,"NativeAmerican","AAPI", "OtherRace")
dat14_re[factorcols] <- lapply(dat14_re[factorcols], factor)
str(dat14_re)
imputed_dat14_re <- missForest(dat14_re[,2:69], parallelize = 'forests', verbose=T)$ximp #LEAVE OUT The AID

write.csv(imputed_dat14_re, "imputed_dat14_re_JMF.csv", row.names = F)
imputed_dat14_re<-read.csv("imputed_dat14_re_JMF.csv")
str(imputed_dat14_re)
#convert back to numeric format
imputed_dat14_re[factorcols] <- sapply(dat14_re[factorcols], as.numeric)

```

Sample weights
```{r}
load("Wave4Weights.rda")
Wave4Weights<-da21600.0031
Wave4Weights_c<-Wave4Weights[c('AID', 'GSWGT4_2')]
write.csv(Wave4Weights_c, "Wave4Weights_c.csv", row.names = F)
all.equal(as.character(Wave4Weights_c$AID), as.character(dat14_re$AID)) #AID all match

```
## preliminary analysis

prepare the data for preliminary analysis
```{r}
dat_prelim<-cbind(dat14["AID"], imputed_dat14_re)
stepdad<-as.numeric(stepdadtot==0)
#then also add some control variables into it
extra_control<-cbind(dat1["AID"], BioDad, stepdad)
dat_prelim_c<-merge(dat_prelim, extra_control, by="AID")
dat_prelim_c[,2:71]<-scale(dat_prelim_c[,2:71]) #scaling all the variables for standardized coefs
describe(dat_prelim_c)
```

```{r}
prelim_coefs<-data.frame("Variable" = names(dat14_re)[c(2:57, 65:68)])
prelim_coefs
prelim_coefs$control<-rep(NA, 60)
prelim_coefs$coefs<-rep(NA,60)
prelim_coefs$CI_L<-rep(NA,60)
prelim_coefs$CI_U<-rep(NA,60)

prelim_coefs[prelim_coefs$Variable%in%c("MotherEducation", "MotherOccupationPrestige", "MotherWorkForPay","MotherRelationshipQuality","MotherSharedActivity","MotherSupervision","MotherInvolveSchool","MotherEducationExpectation","MotherNativity","MotherAlcoholic","MotherObese","MotherDisable"),]$control<-"ResidentMotherPresence"
prelim_coefs[prelim_coefs$Variable%in%c("FatherEducation", "FatherOccupationPrestige", "FatherWorkForPay","FatherRelationshipQuality","FatherSharedActivity","FatherSupervision","FatherInvolveSchool","FatherEducationExpectation","FatherNativity","FatherAlcoholic","FatherObese","FatherDisable"),]$control<-"ResidentFatherPresence"
prelim_coefs[prelim_coefs$Variable%in%c("ParentPartnerRelationship"),]$control<-"ParentPartnered"
prelim_coefs[prelim_coefs$Variable%in%c("NonresidentBioFatherDeceased"),]$control<-"BioDad" 
prelim_coefs[prelim_coefs$Variable%in%c("YearsLiveStepfather"),]$control<-"stepdad" 
```

### LM for individual predictors
```{r}
for(i in 1:60){
  varname<-prelim_coefs$Variable[i]
  control<-prelim_coefs$control[i]
  if(is.na(control)){
    formulai<-paste("EducationAttainment~", varname, "+Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace")
  }
  else{
    formulai<-paste("EducationAttainment~", varname,"+", control, "+Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace")
  }
  print(formulai)
  lm_model<-lm(as.formula(formulai), data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
  prelim_coefs$coefs[i]<-lm_model$coefficients[2]
  prelim_coefs$CI_L[i]<-confint(lm_model, varname, 1-0.05/60)[1]
  prelim_coefs$CI_U[i]<-confint(lm_model, varname, 1-0.05/60)[2]
}

write.csv(prelim_coefs, "prelim_coefs.csv", row.names = F)
```

Visualizing the lm coefficients
```{r}
#first, order the coefficients
prelim_coefs_c<-prelim_coefs[order(prelim_coefs$coefs, decreasing = T),]
colnames(prelim_coefs_c)[3]<-"Coefficient"
varorder<-rev(prelim_coefs_c$Variable)
prelim_coefs_c$Variable<-factor(prelim_coefs_c$Variable, levels= varorder)
library(ggplot2)
p<-ggplot(prelim_coefs_c, 
       aes(x = Variable, y = Coefficient)) +
        geom_hline(yintercept = 0, 
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = Variable, 
                    y = Coefficient)) + 
        geom_linerange(aes(x = Variable, 
                     ymin = CI_L,
                     ymax = CI_U),
                   lwd = 1/2) +
        ggtitle("Outcome: Educational Attainment") +
        coord_flip()
ggsave("prelim_coefs.png", width = 15, height = 30, units = "cm")

write.csv(prelim_coefs_c, "prelim_coefs_c.csv", row.names = F)

```

Count the significant ones
```{r}
prelim_coefs_c$sig<-rep(NA, 60)
prelim_coefs_c$sig<-as.numeric((prelim_coefs_c$CI_L>0 & prelim_coefs_c$CI_U>0)|(prelim_coefs_c$CI_L<0 & prelim_coefs_c$CI_U<0))
sum(prelim_coefs_c$sig==1) #46 variables
```

Rank by absolute values
```{r}
prelim_coefs_c$Coefficient_abs<-abs(prelim_coefs_c$Coefficient)
prelim_coefs_cc<-prelim_coefs_c[order(prelim_coefs_c$Coefficient_abs, decreasing = T),]
```

### R_squared for different domains of family experiences

```{r}
#Demographic variables only
lm_model<-lm(EducationAttainment~Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model) #R2 = 0.0467

#family structural characteristics
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+ ResidentBiologicalParents+ ParentMarried+ ParentPartnered+ HouseholdSize + NumberSiblings+ BirthOrder+ ParentAge+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model) #R2 = 0.1394

#family socioeconomic characteristics
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
               MotherEducation+FatherEducation+MotherOccupationPrestige+FatherOccupationPrestige+ MotherWorkForPay+FatherWorkForPay+FamilyIncome+PublicAssistance+Welfare+EconomicHardship+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = 0.2876

#family relationships and parenting
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+ ParentPartnered+ 
               FamilySocialSupport+SharedDinner+IntergenerationalClosure+ParentConnectionFriends+ MotherRelationshipQuality+FatherRelationshipQuality+MotherSharedActivity+FatherSharedActivity+ ParentPartnerRelationship+ParentControl+MotherSupervision+FatherSupervision+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model) #R2 = 0.1632

#parent involvement with education
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
              MotherInvolveSchool+FatherInvolveSchool+ParentPTA+MotherEducationExpectation+ FatherEducationExpectation+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = 0.1927

#family sociocultural characteristics
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
              MotherNativity+FatherNativity+ParentYearsInUS+ParentReligiosity+EnglishHome+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = 0.1076

#family health resources and behaviors
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
                ParentHealth+ParentSmoke+MotherAlcoholic+FatherAlcoholic+MotherObese+FatherObese+ MotherDisable+FatherDisable+SmokerHousehold+DrugHousehold+FamilyAccessMedicalCare+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = .2031

#parent and family relationship history
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
                ParentPreviousRelationship+NonresidentBioFatherDeceased+BioDad+YearsLiveStepfather+
               stepdad+BornToAdolescentParent+
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = 0.1061

#adverse family experiences
lm_model<-lm(EducationAttainment~ResidentMotherPresence+ ResidentFatherPresence+
                EmotionalAbuse+PhysicalAbuse +BioMotherIncarceration +BioFatherIncarceration +
               Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace, data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model)  #R2 = 0.09413

#ALL
formulai<-paste("EducationAttainment~", paste(prelim_coefs$Variable, collapse = "+"), "+Sex+Age+Latinx+Black+NativeAmerican+AAPI+OtherRace")
lm_model<-lm(as.formula(formulai), data=dat_prelim_c, weights=Wave4Weights_c$GSWGT4_2)
summary(lm_model) #R2 = 0.3714
```

